---
title: "The how and why of R package dependency graphs"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{The how and why of R package dependency graphs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Currently, `depgraph` ships only a single function that plots the dependency graph of an R package based on its `DESCRIPTION` file.
Consider the following example.

```{r fig.align = "cente", fig.cap = "Dependency graph of a historic development version of the R package `multibridge`.", fig.height = 5, fig.width = 7, warning = FALSE}
library("depgraph")

plot_dependency_graph(
  multibridge_pkg
  , suggests = FALSE
  , option = "cividis"
)
```

These plots can be used for at least two purposes:

1. Including dependency graphs in an R package `README` shows users how many packages their work depends on (indirectly).
This is useful as a very rough index of package reliability (less dependencies generally mean less potential for breaking upstream changes) and installation time.
2. Dependency graphs are useful to identify potential to cut down on dependencies.
Briefly, in these graphs you can look for "hot spots" in the network (big bright dots), which represent packages that have many upstream dependencies but are potentially easy to remove because they have few downstream dependencies (that is, only your package depends on them).

The example above has a modest dependency graph; this version of `multibridge` imports `r length(strsplit(multibridge_pkg$imports, ",\n")[[1]])` packages.
Close inspection of the graph shows two illustrative cases:

1. `rlang` is a relatively dark and small node.
This indicates that the package has a lot of downstream dependencies: Several packages in the dependency graph import `rlang`.
If we wanted `multibridge` to not depend on `rlang`, we would have to remove four more packages from the dependency graph (and their respective downstream dependencies).
It seems unlikely that this would be feasible.
Moreover, removing `rlang` from the dependency graph has only a limited impact on the graph because it has no upstream dependencies---it imports no additional packages.
2. `DescTools` is a "hot spot" in the dependency graph:
It has relatively many upstream dependencies.
At the same time no other package in the dependency network imports `DescTools`; it is therefore a promising candidate to reduce the dependency graph.

```{r echo = FALSE}
multibridge_pkg$imports <- gsub("DescTools,\n", "", multibridge_pkg$imports, fixed = TRUE)
```

Inspection of the code (searching the RStudio project for `DescTools` ) revealed that `multibridge` used only one function that could easily be replaced by a custom implementation.
This resulted in a substantial reduction of indirect dependencies, as evident in the new dependency graph.

```{r fig.align = "cente", fig.cap = "Dependency graph of a historic development version of the R package `multibridge` after removing `DescTools` from `Imports`.", fig.height = 5, fig.width = 7, warning = FALSE}
plot_dependency_graph(
  multibridge_pkg
  , suggests = FALSE
  , option = "cividis"
)
```
